Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-27 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
minor changes before submissiion to the LSC

\begin{table}[t]
\begin{tabular}{c | c c}
\hline \hline
Theory & Constraint & This work \\
\hline
EdGB        & $\ell_{\rm EdGB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & -- \\
dCS         & $\ell_{\rm dCS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm dCS} \leqslant 38.7$~km \\
cubic EFT   & -- & $\ell_{\rm cEFT} \leqslant 38.2$~km \\
quartic EFT & $\ell_{\rm qEFT} \leqslant 150$~km~\cite{Sennett:2019bpc}  & $\ell_{\rm qEFT} \leqslant 51.3$~km\hs{$^{\ast}$} \\
\hline \hline
\end{tabular}
\caption{Summary of the upper limits, at $90\%$ credible level,
on the length-scale of the modified gravity theories under investigation.
They were obtained combining Bayesian-inference results from GW150914 and GW200129.
Current constraints on $\ell_{\rm th}$ are also listed.
%
\hs{Our result for the quartic EFT of GR is only in marginal tension with our hard cut-off scale for the
validity of the theory and hence we do still quote it here. See Secs.~\ref{sec:remarks} and~\ref{sec:results_qeft}
for details.}
}
\label{tab:bound_summary}
\end{table}

@@

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with GW
observations of compact binaries where at least one component is a BH.
%
Indeed, Ref.~\cite{Lyu:2022gdr} placed the bound
%
% $\ell_{\rm EdGB} \leqslant 1.33$~km
\hs{$\ell_{\rm EdGB} \leqslant 7.1$~km}
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
% $\ell_{\rm EdGB} \leqslant 1.7$~km,
\hs{$\ell_{\rm EdGB} \leqslant 9.1$~km},
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}\footnote{\hs{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
take into account only the leading-order scalar field interaction arising from the original dilatonic coupling, i.e.,
$f(\varphi) \approx \varphi$ in Eq.~\eqref{eq:action_sgb}.
This results in what is often referred to as shift-symmetric scalar-Gauss-Bonnet theory.
In this theory, NSs do not have scalar monopole charge~\cite{Yagi:2015oca}, while BHs do~\cite{Yunes:2011we,Sotiriou:2013qea,Sotiriou:2014pfa}. Finally, we take note constant $\alpha_{\rm EdGB}$ used in Refs.~\cite{Yagi:2011xp,Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} is related to $\ell_{\rm EdGB}$ as $\ell_{\rm EdGB} = 4 \pi^{1/4} |\alpha_{\rm EdGB}|^{1/2}$.
\hscomm{I had to translate between notations and rewrote the bounds quoted for $\alpha_{\rm EdGB}$ in terms of our $\ell_{\rm EdGB}$.}
}}.
Secondly, the scalar field influences the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant,
results in a coupling between scalar perturbations and gravitational
perturbations of polar parity which, for instance, breaks the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) of Schwarzschild BHs in GR.

Using BH perturbation theory, the QNMs for nonrotating BHs in EdGB gravity
were first computed in Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn}
and were extended \hs{(in the polar sector)} to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.
%
See Ref.~\cite{Bryant:2021xdh} for a study in the geometrical optics limit ($\ell \gg 1$) and
Refs.~\cite{Witek:2018dmd,Okounkova:2019zep} for NR studies.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} =
    \int \dV
    \left[
    \hs{\tfrac{1}{16\pi}} R - \tfrac{1}{2}(\partial \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
The variation of the Pontryagin density is a boundary
term that does not contribute to the field equations in
four-dimensions~\cite{Jackiw:2003pm}.
%
However, the Pontryagin density can modify the field equations when coupled to $\vartheta$;
the strength of the coupling set by $\ell_{\rm dCS}$, with dimensions of length.
%
% However, a coupling with $\vartheta$ can modify the field equations.
% %
% The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
% dimensions of length.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}$    & 0.3737                    \\
$\tau^{(0)}$      & 11.2407                   \\
$\omega^{(1)}$    & 0.1258                    \\
$\tau^{(1)}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}$      & $-$1.1014                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
%
\hscomm{I added the values of $p$ to the column headers to keep both columns with the same number of rows. I also
removed the subscript 220 because we already said we consider only this mode.}
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that they should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final
results using $\varepsilon = 1/2$, \hs{but we stress that there is no
fundamental justification for this choice.}
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %

@@

In Fig.~\ref{fig:dcs_cdf} we show the CDF for GW150914, we see that with $\gm = m_2$,
Eq.~\eqref{eq:eft_bound} is not satisfied unless $\varepsilon \approx 0.9$ (with only $j=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $j=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_{\rm f}$. In this case, we find that
with or without spin corrections Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leqslant 1/2$ (i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}).
%
\hscomm{I commented out the 68\% results for consistency with the other sections.}
%
This means that with our model's assumptions and using the remnant's source mass $M_{\rm f}$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm dCS} \leqslant 41.9~\textrm{km}
\quad \textrm{at 90\% credible level},
\end{equation}
%
on dCS gravity and would constitute \emph{the first bound on this theory with GW
observations alone}.
%
% We also note that since the CDF grows fast in approximate domain $\varepsilon = [0.2, 0.4]$,
% that a stronger, albeit at lower credible level,
% %
% \begin{equation}
% \ell_{\rm dCS} \leqslant 37.4~\textrm{km}
% \quad \textrm{at 68\% credible level},
% \end{equation}
% %
% can also be placed, when $N_{\rm max}=0$.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for both events, calculated in the
same way as already described for the EdGB and dCS theories.
%
We see that curves are very similar to those of dCS gravity for GW200129 (see
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for both events both when
$\mathfrak{m} = M_{\rm f}$, $\varepsilon = 1/2$\hs{, and $N_{\rm max} = 0$.}
%
This allows us to place the combined bound of
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 38.2~\textrm{km}, \quad \textrm{at 90\% credible level}\,.
\end{equation}
%
As also happened for our study for dCS, the find that, for
the cubic EFT, the ParSpec bound is violated by the $N_{\rm max} = 1$ corrections to
the QNMs, meaning that we cannot use this case to draw any meaningful
constraint on this parameter.
%
We summarize our results in Table~\ref{tab:summary_ceft}.

@@

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that while the Parspec bound is satisfied, the EFT bound is only marginally so,
%
As shown in Fig.~\ref{fig:qEFT_cdf}, the 90\% credible level is reached for $\varepsilon \approx 0.58$
(in the case of GW15094) and for $\varepsilon \approx 0.64$ (in the case of GW200129).
%
\hs{Having in mind that the cut off $\varepsilon = 1/2$ is not fundamental, but to keep consistency
across our analysis, our final result
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km},
    %\quad \textrm{at 90\% credible level},
\end{equation}
%
at 90\% credible level should be taken lightly.}
%
However, we \emph{can} claim the validity of the bound above, but at a lower,
68\% credible level.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold_corrected.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at \hs{68\% credibility}.
}
\label{fig:qEFT_cdf}
\end{figure}

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{\rm f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
% \hline
%   & GW150914 &  &     &    \\
% 1 & GW200129 &  &     &    \\
%   & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$, and $N_{\rm max} = 0$. The quoted
result correspond \hs{to 90\% credible values \emph{if} we allow for a more flexible cut-off $\varepsilon \lesssim 0.65$.
However, the result is robust for the cut off $\varepsilon = 1/2$, at 65\% credible level.}
}
\label{tab:summary_qeft}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), in the particular case of $\lame = \lamo = 1$.
%
%
We then linearize the resulting expression in $\chi$ and consider $m=2$ the harmonic.
%
\hscomm{Instead calling them ``$+$'' and ``$-$'' following the notation used in~\cite{Cano:2021myl},
I call them axial/polar to follow our conventions.}
%
As an outcome, we find that the fundamental axial-parity QNM is the least damped one, and it is
the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12,
\end{align}
%
for this theory.

>>>>

\begin{table}[t]
\begin{tabular}{c | c c}
\hline \hline
Theory & Constraint & This work \\
\hline
EdGB        & $\ell_{\rm EdGB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & -- \\
dCS         & $\ell_{\rm dCS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm dCS} \leqslant 38.7$~km \\
cubic EFT   & -- & $\ell_{\rm cEFT} \leqslant 38.2$~km \\
quartic EFT & $\ell_{\rm qEFT} \leqslant 150$~km~\cite{Sennett:2019bpc}  & $\ell_{\rm qEFT} \leqslant 51.3$~km~\footnote{We notice that our result for the quartic EFT of GR is only in marginal tension with our hard cut-off scale for the
validity of the theory, and hence we do still quote it here (see Secs.~\ref{sec:remarks} and~\ref{sec:results_qeft}
for details).}
\\
\hline \hline
\end{tabular}
\caption{Summary of the upper limits, at $90\%$ credible level,
on the length-scale of the modified gravity theories under investigation.
They were obtained combining Bayesian-inference results from GW150914 and GW200129.
Current constraints on $\ell_{\rm th}$ are also listed.}
\label{tab:bound_summary}
\end{table}

@@

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with GW
observations of compact binaries where at least one component is a BH.
%
Indeed, Ref.~\cite{Lyu:2022gdr} placed the bound
%
% $\ell_{\rm EdGB} \leqslant 1.33$~km
$\ell_{\rm EdGB} \leqslant 7.1$~km
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
% $\ell_{\rm EdGB} \leqslant 1.7$~km,
$\ell_{\rm EdGB} \leqslant 9.1$~km,
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}\footnote{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
they take into account only the leading-order scalar field interaction arising from the original dilatonic coupling 
(i.e., $f(\varphi) \approx \varphi$ in Eq.~\eqref{eq:action_sgb}).
This results in what is often referred to as shift-symmetric scalar-Gauss-Bonnet theory.
In this theory, NSs do not have scalar monopole charge~\cite{Yagi:2015oca}, while BHs do~\cite{Yunes:2011we,Sotiriou:2013qea,Sotiriou:2014pfa}. Finally, we note that the constant $\alpha_{\rm EdGB}$ used in Refs.~\cite{Yagi:2011xp,Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} is related to $\ell_{\rm EdGB}$ as $\ell_{\rm EdGB} = 4 \pi^{1/4} |\alpha_{\rm EdGB}|^{1/2}$.}.
Secondly, the scalar field influences the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant,
results in a coupling between scalar perturbations and gravitational
perturbations of polar parity which, for instance, breaks the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) of Schwarzschild BHs in GR.

Using BH perturbation theory, the QNMs for nonrotating BHs in EdGB gravity
were first computed in Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn}
and were extended (in the polar sector) to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.
%
See Ref.~\cite{Bryant:2021xdh} for a study in the geometrical optics limit ($\ell \gg 1$) and
Refs.~\cite{Witek:2018dmd,Okounkova:2019zep} for NR studies.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} =
    \int \dV
    \left[
    \tfrac{1}{16\pi}\, R - \tfrac{1}{2}(\partial \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
The variation of the Pontryagin density is a boundary
term that does not contribute to the field equations in
four-dimensions~\cite{Jackiw:2003pm}.
%
However, the Pontryagin density can modify the field equations when coupled to $\vartheta$;
the strength of the coupling set by $\ell_{\rm dCS}$, with dimensions of length.
%
% However, a coupling with $\vartheta$ can modify the field equations.
% %
% The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
% dimensions of length.

@@

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}$    & 0.3737                    \\
$\tau^{(0)}$      & 11.2407                   \\
$\omega^{(1)}$    & 0.1258                    \\
$\tau^{(1)}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}$      & $-$1.1014                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we consider together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also include for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
%
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that they should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final
results using $\varepsilon = 1/2$, but we stress that there is no
fundamental justification for this choice.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %

@@

In Fig.~\ref{fig:dcs_cdf} we show the CDF for GW150914, we see that with $\gm = m_2$,
Eq.~\eqref{eq:eft_bound} is not satisfied unless $\varepsilon \approx 0.9$ (with only $j=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $j=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_{\rm f}$. In this case, we find that
with or without spin corrections Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leqslant 1/2$ (i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}).
%
This means that with our model's assumptions and using the remnant's source mass $M_{\rm f}$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm dCS} \leqslant 41.9~\textrm{km}
\quad \textrm{at 90\% credible level},
\end{equation}
%
on dCS gravity and would constitute \emph{the first bound on this theory with GW
observations alone}.
%
% We also note that since the CDF grows fast in approximate domain $\varepsilon = [0.2, 0.4]$,
% that a stronger, albeit at lower credible level,
% %
% \begin{equation}
% \ell_{\rm dCS} \leqslant 37.4~\textrm{km}
% \quad \textrm{at 68\% credible level},
% \end{equation}
% %
% can also be placed, when $N_{\rm max}=0$.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for both events, calculated in the
same way as already described for the EdGB and dCS theories.
%
We see that curves are very similar to those of dCS gravity for GW200129 (see
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for both events both when
$\mathfrak{m} = M_{\rm f}$, $\varepsilon = 1/2$, and $N_{\rm max} = 0$.
%
This allows us to place the combined bound of
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 38.2~\textrm{km}, \quad \textrm{at 90\% credible level}\,.
\end{equation}
%
As also happened for our study for dCS, the find that, for
the cubic EFT, the ParSpec bound is violated by the $N_{\rm max} = 1$ corrections to
the QNMs, meaning that we cannot use this case to draw any meaningful
constraint on this parameter.
%
We summarize our results in Table~\ref{tab:summary_ceft}.

@@

Let us now consider the quartic EFT of GR, as our final example.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR.
%
We find that while the Parspec bound is satisfied, the EFT bound is only marginally so,
%
As shown in Fig.~\ref{fig:qEFT_cdf}, the 90\% credible level is reached for $\varepsilon \approx 0.58$
(in the case of GW15094) and for $\varepsilon \approx 0.64$ (in the case of GW200129).
%
Having in mind that the cut off $\varepsilon = 1/2$ is not fundamental, but to keep consistency
across our analysis, our final result
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km},
    %\quad \textrm{at 90\% credible level},
\end{equation}
%
at 90\% credible level should be taken lightly.
%
However, we \emph{can} claim the validity of the bound above, but at a lower,
68\% credible level.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold_corrected.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at 68\% credibility.
}
\label{fig:qEFT_cdf}
\end{figure}

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{\rm f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
% \hline
%   & GW150914 &  &     &    \\
% 1 & GW200129 &  &     &    \\
%   & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$, and $N_{\rm max} = 0$. The quoted
result correspond to 90\% credible values \emph{if} we allow for a more flexible cut-off $\varepsilon \lesssim 0.65$.
However, the result is robust for the cut off $\varepsilon = 1/2$, at 65\% credible level.
}
\label{tab:summary_qeft}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), in the particular case of $\lame = \lamo = 1$.
%
%
We then linearize the resulting expression in $\chi$ and consider $m=2$ the harmonic.
%
As an outcome, we find that the fundamental axial-parity QNM is the least damped one, and it is
the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12,
\end{align}
%
for this theory.
